---
title: "BSMM-lab-3"
subtitle: "BSMM 8740 Fall 2023"
author: "Archana Vijayakumar Sreekala"
date: "October 12th, 2023"
format: html
editor: visual
self-contained: true
---

## Setup

Load packages and data:

```{r load-pkg-data}
#| message: false

library(magrittr) # the pipe 
library(tidyverse) # for data wrangling + visualization 
library(tidymodels) # for modeling 
library(gt) # for pretty tables 
library(ggplot2) # to draw the plots
library(forcats) # to deal with category variables


boston_cocktails <- readr::read_csv('data/boston_cocktails.csv')
```

## Exercises

### Exercise 1

```{r #To assess the quality of the dataset}

view (boston_cocktails) # view the dataset 
skimr::skim(boston_cocktails) # skimr to view the dataset quality 

# Data Explorer introduce to view the details of missing and col info 
DataExplorer::introduce(boston_cocktails) 

# To take summary and find mean of measure value 

median_measure_number <- boston_cocktails %>% summarise(median_measure_number=median(measure_number)) 
median_measure_number

# To display the answer 
stringr::str_glue( "The median measure amount across across all cocktails is {median_measure_number}.")
```

### Exercise 2

```{r #o pivot the table and find the gin amount in a cocktail}
  
# To form the pivot table by selecting the required fields 

pivot_table <- boston_cocktails %>% select(name, category, ingredient, measure_number) %>%     pivot_wider(names_from = ingredient, values_from = measure_number, values_fill = 0) 

# To clean the names using janitor 

clean_col <- pivot_table %>% janitor::clean_names() 

# To drop the NA columns 

clean_col_na <- clean_col %>% drop_na() 

# To filter based on "Leap Frog Highball" and find the amount of gin 

gin_amount <- clean_col_na %>% dplyr::filter(name == "Leap Frog Highball") %>% dplyr::select(gin) 

stringr::str_glue( "The Leap Frog Highball contains {gin_amount} oz of gin")
```

### Exercise 3

```{r #To implement recipe and arrive at the no of predictor variables}

library(recipes) #load library for recipes

# create recipe, update roles, normalise and perform pca
recipe_var <- recipe(~ ., data =clean_col_na) %>%
              update_role(name, category, new_role = "id") %>% 
              recipes::step_normalize(all_numeric()) %>%
              step_pca(all_numeric())

# prep the final data and assign to a variable
recipe_prep_var <- prep(recipe_var)

recipe_prep_var
```

40 predictor variables are prepped by the recipe.

### Exercise 4

```{r}

# To perform tidy on the prepped data

recipe_tidy <- recipes::tidy(recipe_prep_var)
recipe_tidy

# To display the normalization data 

recipe_norm <- recipe_prep_var %>% recipes::tidy(1)
recipe_norm

# To find the most used ingredient on average based on statistic as "mean"
Most_used_ingr_norm_mean <- recipe_norm %>% dplyr::filter(statistic == "mean") %>%
                dplyr::arrange(desc(value)) %>% slice_head(n=1)

Most_used_ingr_norm_mean

# To find the most used ingredient on average based on statistic as "sd"
Most_used_ingr_norm_sd <- recipe_norm %>% dplyr::filter(statistic == "sd") %>%
                dplyr::arrange(desc(value)) %>% slice_head(n=1)

Most_used_ingr_norm_sd

# To find the most used ingredient on average without normalization method

most_used_ingredient <- boston_cocktails %>%
              group_by(ingredient) %>%
              summarise(average_count = mean(measure_number)) %>%
              arrange(desc(average_count))  
most_used_ingredient
```

On average the most used ingredient in the Boston Cocktails dataset is **gin** based on **normalised data with statistic = 'mean'.**

On average the most used ingredient in the Boston Cocktails dataset is **Orange juice** based on **normalised data with statistic = 'sd'**

On average the most used ingredient in the Boston Cocktails dataset is **cranberry juice** based on **with out normalised data.**

### Exercise 5

```{r}

# To get the data for pca
recipe_pca <- recipe_prep_var %>% recipes::tidy(2)
view(recipe_pca)

# To filter the pca values from 1 to 5 and mutate to factor data and order it
recipe_pca_1_5 <- recipe_pca %>% dplyr::filter(component %in% c("PC1","PC2","PC3","PC4","PC5")) %>% 
                                 dplyr::mutate(order = forcats::fct_inorder(component))  

recipe_pca_1_5

# To plot the values using ggplot 
recipe_pca_1_5 %>% ggplot(aes(value, terms, fill = terms)) +
                   geom_col(show.legend = FALSE) +
                   facet_wrap(~component, nrow = 1) +
                   labs(y = NULL) +
                   theme(axis.text=element_text(size=7),
                   axis.title=element_text(size=14,face="bold"))

```

Describe the drinks represented by PC1?

The drink is highly syrupy in nature, very less in sugary flavor and less calories because of that. This is a drink having moderate alcohol content with Tequila as the main ingredient along with juice such as lime juice.

### Exercise 6

```{r}

# Filter data for pca from PC1 to PC4 and get top 8 ingredients and plot
recipe_pca %>% filter(component %in% c("PC1", "PC2", "PC3", "PC4")) %>%  #filter based on PC1 to PC4
               group_by(component) %>%
               slice_max(order_by = abs(value), n = 8) %>%               #slice the absvalue of first 8 ingredients
               gt() %>%                                                  # use gt() and color the data based on value
               data_color(
                 columns = c(value),
                 colors = scales::col_numeric(
                 palette = c("red", "green"),
                 domain = c(-1, 1)
                 )
                ) 
```

The characteristic alcoholic beverage of each of the first 4 principle components as below:

**PC1:**

The main noticeable ingredients with the highest absolute values: simple syrup (0.3125978) and powdered sugar (-0.4764442).

Characteristic alcoholic beverage of PC1 shows that, it is a syrupy textured mixed drink (with lime juice and **Tequila**), having less sugar content.

The most used alcoholic beverage: Tequila.

**PC2:**

The main noticeable ingredients with the highest absolute values: dry **vermouth** (0.4332412), sweet **vermouth** (0.3754366), **gin** (0.2437374)

Characteristic alcoholic beverage of PC2 shows that, this drink is in a cocktail category with the main ingredient as vermouth with a mix of gin, and less sugary content.

The most used alcoholic beverage: gin, vermouth.

**PC3:**

Ingredients with the highest absolute values: **gin** (0.3780348), egg white (0.3094631), followed by lemon juice(0.2318927) and apricot flavored brandy (0.2134899).

Characteristic alcoholic beverage of PC3 shows that, it is likely associated with a fruity mix of cocktail that has a prominent content of gin and also include egg white.

The most used alcoholic beverage: gin.

**PC4:**

Ingredients with the highest absolute values: grenadine (0.4068737), orange juice (0.3625622), **vodka**(0.3008406).

Characteristic alcoholic beverage of PC4 shows that, it is also a cocktail with mostly a wide variety mix of fruity flavors (pomegranate, orange, cranberry), cream with vodka.

The most used alcoholic beverage: vodka.

### Exercise 7

How would you interpret the results of a PCA analysis for a client?

```{r}

# To bake the prepped data and plot using ggplot for PC1 and PC2

recipes::bake(recipe_prep_var,clean_col) %>% 
          ggplot(aes(PC1, PC2, label = name)) +
          geom_point(aes(color = category), alpha = 0.7, size = 2) +
          geom_text(check_overlap = TRUE, hjust = "inward") + 
          labs(color = NULL)

```

Based on the plot data, it shows that PC1 set is having a lesser affinity towards cocktail types such as Cocktail classics, cordial and Liquers and they are having moderate affinity towards the varieties such as Gin and a high affinity towards the strong alcoholic varieties such as Rum, Tequila, Vodka, Whiskies.

This shows that PC1 is more into strong alcoholic varieities than the mixed cocktail non alcoholic types.

For PC2, it shows that the cocktail variations carry a major part, but based on the type of cocktail, the affinity varies. That is, some cocktails in classic cocktails or in the type: cordial and Liquers are least preferred such as Lemonade, Coffee flip etc, where as some others such as Sphinx cocktail, Nineteenth Hole etc are highly preferred. In the case of alcoholic categories such as Vodka, Gin, Tequila etc are moderately preferred.

This shows that PC2 is more in to certain variations of cocktails (not all types) and they moderately prefer alcoholic varieties.
