---
title: "BSMM-lab-3"
subtitle: "BSMM 8740 Fall 2023"
author: "Archana Vijayakumar Sreekala"
date: "October 12th, 2023"
format: html
editor: visual
self-contained: true
---

## Setup

Load packages and data:

```{r load-pkg-data}
#| message: false

library(magrittr) # the pipe 
library(tidyverse) # for data wrangling + visualization 
library(tidymodels) # for modeling 
library(gt) # for pretty tables 

boston_cocktails <- readr::read_csv('data/boston_cocktails.csv')
```

## Exercises

### Exercise 1

```{r #To assess the quality of the dataset}

view (boston_cocktails) # view the dataset 
skimr::skim(boston_cocktails) # skimr to view the dataset quality 

# Data Explorer introduce to view the details of missing and col info 
DataExplorer::introduce(boston_cocktails) 

# To take summary and find mean of measure value 

median_measure_number <- boston_cocktails %>% summarise(median_measure_number=median(measure_number)) 
median_measure_number

# To display the answer 
stringr::str_glue( "The median measure amount across across all cocktails is {median_measure_number}.")
```

### Exercise 2

```{r #o pivot the table and find the gin amount in a cocktail}
  
# To form the pivot table by selecting the required fields 

pivot_table <- boston_cocktails %>% select(name, category, ingredient, measure_number) %>%     pivot_wider(names_from = ingredient, values_from = measure_number, values_fill = 0) 

# To clean the names using janitor 

clean_col <- pivot_table %>% janitor::clean_names() 

# To drop the NA columns 

clean_col_na <- clean_col %>% drop_na() 

# To filter based on "Leap Frog Highball" and find the amount of gin 

gin_amount <- clean_col_na %>% dplyr::filter(name == "Leap Frog Highball") %>% dplyr::select(gin) 

stringr::str_glue( "The Leap Frog Highball contains {gin_amount} oz of gin")
```

### Exercise 3

\_\_\_ predictor variables are prepped by the recipe.

### Exercise 4

On average the most used ingredient in the Boston Cocktails dataset is \_\_\_\_.

### Exercise 5

Describe describe the drinks represented by PC1?

### Exercise 6

The characteristic alcoholic beverage of each of the first 4 principle components is \_\_\_.

### Exercise 7

How would you interpret the results of a PCA analysis for a client?
